---
title: "CIBERSORTx_Analyses"
output: html_notebook
---
This document shows all the different pipelines followed to create the result plots found in the thesis from the CIBERSORTx data

# Libraries Needed
```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(paletteer)
library(rcartocolor)
library(pheatmap)
library(gridExtra)
```

# Setting the colors that will be needed for each SM
For LM22:
```{r}
stackcolors <- c(
  "B.cells.naive" = "yellow",  
  "B.cells.memory" = "gold",  
  "Plasma.cells" = "orange",    
  "T.cells.CD8" = "lightgreen",    
  "T.cells.CD4.naive" = "darkgreen",
  "T.cells.CD4.memory.resting" = "darkolivegreen",  
  "T.cells.CD4.memory.activated" = "darkolivegreen1", 
  "T.cells.follicular.helper" = "darkolivegreen3",  
  "T.cells.regulatory..Tregs." = "darkolivegreen4", 
  "T.cells.gamma.delta" = "chartreuse4",  
  "NK.cells.resting" = "#3B82F6",  
  "NK.cells.activated" = "#1E3A8A",  
  "Monocytes" = "darkorchid1",  
  "Macrophages.M0" = "darkmagenta",  
  "Macrophages.M1" = "brown1",  
  "Macrophages.M2" = "brown3",  
  "Dendritic.cells.resting" = "darkred",  
  "Dendritic.cells.activated" = "#1ABC9C",  
  "Mast.cells.resting" = "lightgrey",  
  "Mast.cells.activated" = "grey",  
  "Eosinophils" = "darkgrey",  
  "Neutrophils" = "gray28"
)
```
For Van der Wijst SM:
```{r}
stackcolors2 <- c(
  "B" = "yellow",  
  "Plasma" = "orange",    
  "CD8_T" = "lightgreen",    
  "CD4_T" = "darkgreen",
  "CD56bright_NK" = "#3B82F6",  
  "CD56dim_NK" = "#1E3A8A",  
  "cMonocyte" = "darkorchid1",
  "ncMonocyte" = "darkorchid4",
  "Megakaryocyte" = "burlywood",
  "pDC" = "darkred",  
  "mDC" = "#1ABC9C"
)
```
For Monaco:
```{r}
stackcolors3 <- c(
  "B.Memory" = "gold",      
  "B.Naive" = "yellow",
  "Basophils.LD"= "darkgrey",
  "MAIT" = "darkolivegreen3",          
  "Monocytes.C" = "darkorchid1",   
  "Monocytes.NC.I" = "darkorchid4",
  "NK" = "#1E3A8A",             
  "Neutrophils.LD" ="gray28", 
  "Plasmablasts" = "orange",   
  "T.CD4.Memory" =  "darkolivegreen",  
  "T.CD4.Naive" =  "darkgreen",   
  "T.CD8.Memory" = "darkolivegreen1",
  "T.CD8.Naive"  = "lightgreen",  
  "T.gd.Vd2" = "chartreuse4",
  "T.gd.non.Vd2" = "chartreuse2",
  "mDCs"  = "#1ABC9C",         
  "pDCs" = "darkred"  
)
```

# Transcriptome Matrix 1 vs LM22
*Stack bar plot*
```{r}
TM1vsLM22_d23_raw <- read.csv("CIBERSORTx_Results.csv")

TM1vsLM22_d23 <- TM1vsLM22_d23_raw[,-c(24,25,26)]

TM1vsLM22_d23 <- TM1vsLM22_d23 %>% pivot_longer(!Mixture, names_to = "cell_types", values_to = "cell_fraction")

TM1vsLM22_d23 <- TM1vsLM22_d23 %>% rename(patients = Mixture)

LM22order <- c("Neutrophils", "Eosinophils", "Mast.cells.resting", "Mast.cells.activated", "Dendritic.cells.resting", "Dendritic.cells.activated", "Macrophages.M2","Macrophages.M1", "Macrophages.M0", "Monocytes", "NK.cells.activated","NK.cells.resting", "T.cells.gamma.delta", "T.cells.regulatory..Tregs.", "T.cells.follicular.helper", "T.cells.CD4.memory.activated", "T.cells.CD4.memory.resting", "T.cells.CD4.naive", "T.cells.CD8", "Plasma.cells", "B.cells.memory", "B.cells.naive")
TM1vsLM22_d23$cell_types <- factor(TM1vsLM22_d23$cell_types, levels = LM22order)

ggplot(data = TM1vsLM22_d23, mapping = aes(x = patients, y = cell_fraction * 100, fill = cell_types)) + 
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_manual(values = stackcolors) +
  scale_y_continuous(breaks = seq(0, 100, by = 10)) +
  labs(y = "Relative Percent", fill= "Leukocyte cell types") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size= 5), axis.title.x = element_text(size = 9), axis.title.y = element_text(size = 9),legend.text = element_text(size = 7), legend.title =element_text(size = 9), legend.key.size = unit(0.8, "lines"))
```

*Scatter plot*
```{r}
stats_TM1_LM22 <- TM1vsLM22_d23 %>%
  group_by(cell_types) %>%
  summarise(
    mean_fraction = mean(cell_fraction, na.rm = TRUE),
    sd_fraction = sd(cell_fraction, na.rm = TRUE)
  )

stats_TM1_LM22.log <- stats_TM1_LM22 %>%
  mutate(
   log2_mean_fraction = log2(mean_fraction + 1),
    log2_sd_fraction = log2(sd_fraction + 1)
  )

LM22 <- ggplot(stats_TM1_LM22.log, aes(x = log2_mean_fraction, y = log2_sd_fraction, color = cell_types, label = cell_types)) +
  geom_point(size = 3) +
  geom_text(vjust = -0.5, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red",linetype = "dashed") +
  scale_color_manual(name = "Cell Types", values = stackcolors) +
  coord_cartesian(xlim = c(-0.06, 0.6), ylim = c(0, 0.20)) +
  theme_classic() +
  labs(
    title = "Cell type representation of Transcriptome Matrix 1",
    x = "Log2(Mean Cell Fraction +1)",
    y = "Log2(Standard Deviation +1)",
    caption = "(Based on LM22 Signature Matrix)"
  ) +
  theme(
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.3, "cm")
  )
print(LM22)
```

# Transcriptome Matrix 1 vs Van der Wijst
*Stack bar plot*
```{r}
TM1vsVdW_raw <- read.csv("CIBERSORTx_Results.csv")

TM1vsVdW <- TM1vsVdW_raw [,-c(13,14,15)]
TM1vsVdW <- TM1vsVdW  %>% pivot_longer(!Mixture, names_to = "cell_types", values_to = "cell_fraction")
TM1vsVdW <- TM1vsVdW  %>% rename(patients = Mixture)

TM1inVdW <- TM1vsVdW %>%
  group_by(cell_types) %>%
  summarise(mean_fraction = mean(cell_fraction, na.rm = TRUE)) %>%
  arrange(mean_fraction) %>%
  pull(cell_types)

TM1vsVdW$cell_types <- factor(TM1vsVdW$cell_types, levels = TM1inVdW)


ggplot(data = TM1vsVdW, mapping = aes(x = patients, y = cell_fraction * 100, fill = cell_types)) + 
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_manual(values = stackcolors2) +
  scale_y_continuous(breaks = seq(0, 100, by = 10)) +
  labs(y = "Relative Percent", fill= "Leukocyte cell types") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size= 5), axis.title.x = element_text(size = 9), axis.title.y = element_text(size = 9),legend.text = element_text(size = 7), legend.title =element_text(size = 9), legend.key.size = unit(0.8, "lines"))
```

*Scatter plot*
```{r}
stats_TM1_VdW <- TM1vsVdW %>%
  group_by(cell_types) %>%
  summarise(
    mean_fraction = mean(cell_fraction, na.rm = TRUE),
    sd_fraction = sd(cell_fraction, na.rm = TRUE)
  )

stats_TM1_VdW.log <- stats_TM1_VdW %>%
  mutate(
   log2_mean_fraction = log2(mean_fraction + 1),
    log2_sd_fraction = log2(sd_fraction + 1)
  )

VdW <- ggplot(stats_TM1_VdW.log, aes(x = log2_mean_fraction, y = log2_sd_fraction, color = cell_types, label = cell_types)) +
  geom_point(size = 3) +
  geom_text(vjust = -0.5, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red",linetype = "dashed") +
  scale_color_manual(name = "Cell Types", values = stackcolors2) +
  coord_cartesian(xlim =  c(-0.06, 0.6), ylim = c(0, 0.20)) + 
  theme_classic() +
  labs(
    title = "Cell type representation of Transcriptome Matrix 1",
    x = "Log2(Mean Cell Fraction +1)",
    y = "Log2(Standard Deviation +1)",
    caption = "(Based on Van der Wijst Signature Matrix)"
  ) +
  theme(
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.3, "cm")
  )
print(VdW)
```

# Transcriptome Matrix 1 vs default SM from ABIS 
```{r}
TM1_ABIS <- read.csv("CIBERSORTx_Results.csv")

TM1_ABIS <- TM1_ABIS[, 1:(ncol(TM1_ABIS)-3)]  
TM1_ABIS <- TM1_ABIS %>% pivot_longer(!Mixture, names_to = "cell_types", values_to = "cell_fraction")
TM1_ABIS <- TM1_ABIS %>% rename(patients = Mixture)

stats_TM1_ABIS <- TM1_ABIS %>%
  group_by(cell_types) %>%
  summarise(
    mean_fraction = mean(cell_fraction, na.rm = TRUE),
    sd_fraction = sd(cell_fraction, na.rm = TRUE)
  ) %>%
  mutate(
    log2_mean_fraction = log2(mean_fraction + 1),
    log2_sd_fraction = log2(sd_fraction + 1)
  )

TM1_ABIS <- ggplot(stats_TM1_ABIS, aes(x = log2_mean_fraction, y = log2_sd_fraction, color = cell_types, label = cell_types)) +
  geom_point(size = 3) +
  geom_text(vjust = -0.5, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red",linetype = "dashed") +
  scale_color_manual(name = "Cell Types", values = stackcolors3) +
  coord_cartesian(xlim = c(-0.06, 0.6), ylim = c(0, 0.20)) +  
  theme_classic() +
  labs(
    title = "Cell type representation of Transcriptome Matrix 1",
    x = "Log2(Mean Cell Fraction +1)",
    y = "Log2(Standard Deviation +1)",
    caption = "(Based on ABIS default SM)"
  ) +
  theme(
    legend.text = element_text(size = 5),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.3, "cm"),
    plot.title = element_text(size = 10)
  )
print(TM1_ABIS)
```

# Transcriptome Matrix 2 vs LM22
```{r}
TM2vsLM22_raw <- read.csv("CIBERSORTx_Results.csv")

TM2vsLM22 <- TM2vsLM22_raw[,-c(24,25,26)]

TM2vsLM22  <- TM2vsLM22  %>% pivot_longer(!Mixture, names_to = "cell_types", values_to = "cell_fraction")

TM2vsLM22  <- TM2vsLM22  %>% rename(patients = Mixture)

TM2vsLM22$cell_types <- factor(TM2vsLM22$cell_types, levels = LM22order)

ggplot(data = TM2vsLM22, mapping = aes(x = patients, y = cell_fraction * 100, fill = cell_types)) + 
  geom_bar(position = "stack", stat = "identity") +
  scale_fill_manual(values = stackcolors) +
  scale_y_continuous(breaks = seq(0, 100, by = 10)) +
  labs(y = "Relative Percent", fill= "Leukocyte cell types") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size= 5), axis.title.x = element_text(size = 9), axis.title.y = element_text(size = 9),legend.text = element_text(size = 7), legend.title =element_text(size = 9), legend.key.size = unit(0.8, "lines"))

stats_TM2_LM22 <- TM2vsLM22 %>%
  group_by(cell_types) %>%
  summarise(
    mean_fraction = mean(cell_fraction, na.rm = TRUE),
    sd_fraction = sd(cell_fraction, na.rm = TRUE)
  )

stats_TM2_LM22.log <- stats_TM2_LM22 %>%
  mutate(
   log2_mean_fraction = log2(mean_fraction + 1),
    log2_sd_fraction = log2(sd_fraction + 1)
  )

p4 <- ggplot(stats_TM2_LM22.log, aes(x = log2_mean_fraction, y = log2_sd_fraction, color = cell_types, label = cell_types)) +
  geom_point(size = 3) +
  geom_text(vjust = -0.5, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red",linetype = "dashed") +
  scale_color_manual(name = "Cell Types", values = stackcolors) +
  coord_cartesian(xlim = c(-0.15, 0.6), ylim = c(0, 0.3)) +
  theme_classic() +
  labs(
    title = "Cell type representation of Transcriptome Matrix 2",
    x = "Log2(Mean Cell Fraction +1)",
    y = "Log2(Standard Deviation +1)",
    caption = "(Based on LM22 Signature Matrix)"
  ) +
  theme(
    legend.text = element_text(size = 6),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.3, "cm")
  )
print(p4)
```

# LM22 VS Different conditions
Starting with Malignancy 1 of Brucellosis patients with complete remission after treatment
```{r}
Malignancy1 <- read.csv("CIBERSORTx_Results.csv")

Malignancy1 <- Malignancy1[, 1:(ncol(Malignancy1)-3)]  

Malignancy1 <- Malignancy1 %>% pivot_longer(!Mixture, names_to = "cell_types", values_to = "cell_fraction")

Malignancy1 <- Malignancy1 %>% rename(patients = Mixture)

stats_Malignancy1 <- Malignancy1 %>%
  group_by(cell_types) %>%
  summarise(
    mean_fraction = mean(cell_fraction, na.rm = TRUE),
    sd_fraction = sd(cell_fraction, na.rm = TRUE)
  ) %>%
  mutate(
    log2_mean_fraction = log2(mean_fraction + 1),
    log2_sd_fraction = log2(sd_fraction + 1)
  )

M1 <- ggplot(stats_Malignancy1, aes(x = log2_mean_fraction, y = log2_sd_fraction, color = cell_types, label = cell_types)) +
  geom_point(size = 3) +
  geom_text(vjust = -0.5, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red",linetype = "dashed") +
  scale_color_manual(name = "Cell Types", values = stackcolors) +
  coord_cartesian(xlim = c(-0.15, 0.6), ylim = c(0, 0.15)) +
  theme_classic() +
  labs(
    title = "Cell type representation of Brucellosis patients with complete remission after treatment",
    x = "Log2(Mean Cell Fraction +1)",
    y = "Log2(Standard Deviation +1)",
    caption = "(Based on LM22 Signature Matrix)"
  ) +
  theme(
    legend.text = element_text(size = 5),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.3, "cm"),
    plot.title = element_text(size = 10)
  )
print(M1)
```

Now Malignancy 2 wich is Brucellosis patients with reinfection after treatment
```{r}
Malignancy2 <- read.csv("/CIBERSORTx_Results.csv")

Malignancy2 <- Malignancy2[, 1:(ncol(Malignancy2)-3)]  

Malignancy2 <- Malignancy2 %>% pivot_longer(!Mixture, names_to = "cell_types", values_to = "cell_fraction")

Malignancy2 <- Malignancy2 %>% rename(patients = Mixture)

stats_Malignancy2 <- Malignancy2 %>%
  group_by(cell_types) %>%
  summarise(
    mean_fraction = mean(cell_fraction, na.rm = TRUE),
    sd_fraction = sd(cell_fraction, na.rm = TRUE)
  ) %>%
  mutate(
    log2_mean_fraction = log2(mean_fraction + 1),
    log2_sd_fraction = log2(sd_fraction + 1)
  )

M2 <- ggplot(stats_Malignancy2, aes(x = log2_mean_fraction, y = log2_sd_fraction, color = cell_types, label = cell_types)) +
  geom_point(size = 3) +
  geom_text(vjust = -0.5, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red",linetype = "dashed") +
  scale_color_manual(name = "Cell Types", values = stackcolors) +
  coord_cartesian(xlim = c(-0.15, 0.6), ylim = c(0, 0.15)) +
  theme_classic() +
  labs(
    title = "Cell type representation of Brucellosis patients with reinfection after treatment",
    x = "Log2(Mean Cell Fraction +1)",
    y = "Log2(Standard Deviation +1)",
    caption = "(Based on LM22 Signature Matrix)"
  ) +
  theme(
    legend.text = element_text(size = 5),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.3, "cm"),
    plot.title = element_text(size = 10)
  )
print(M2)
```

Now Malignancy 3 wich is Brucellosis patients with relapse after treatment
```{r}
Malignancy3 <- read.csv("CIBERSORTx_Results.csv")

Malignancy3 <- Malignancy3[, 1:(ncol(Malignancy3)-3)]  

Malignancy3 <- Malignancy3 %>% pivot_longer(!Mixture, names_to = "cell_types", values_to = "cell_fraction")

Malignancy3 <- Malignancy3 %>% rename(patients = Mixture)

stats_Malignancy3 <- Malignancy3 %>%
  group_by(cell_types) %>%
  summarise(
    mean_fraction = mean(cell_fraction, na.rm = TRUE),
    sd_fraction = sd(cell_fraction, na.rm = TRUE)
  ) %>%
  mutate(
    log2_mean_fraction = log2(mean_fraction + 1),
    log2_sd_fraction = log2(sd_fraction + 1)
  )

M3 <- ggplot(stats_Malignancy3, aes(x = log2_mean_fraction, y = log2_sd_fraction, color = cell_types, label = cell_types)) +
  geom_point(size = 3) +
  geom_text(vjust = -0.5, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red",linetype = "dashed") +
  scale_color_manual(name = "Cell Types", values = stackcolors) +
  coord_cartesian(xlim = c(-0.15, 0.6), ylim = c(0, 0.15)) +
  theme_classic() +
  labs(
    title = "Cell type representation of Brucellosis patients with relapse after treatment",
    x = "Log2(Mean Cell Fraction +1)",
    y = "Log2(Standard Deviation +1)",
    caption = "(Based on LM22 Signature Matrix)"
  ) +
  theme(
    legend.text = element_text(size = 5),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.3, "cm"),
    plot.title = element_text(size = 10)
  )
print(M3)
```

Now Malignancy 4 wich is Lehismaniosis patients
```{r}
Malignancy4 <- read.csv("CIBERSORTx_Results.csv")

Malignancy4 <- Malignancy4[, 1:(ncol(Malignancy4)-3)]  
Malignancy4 <- Malignancy4 %>% pivot_longer(!Mixture, names_to = "cell_types", values_to = "cell_fraction")
Malignancy4 <- Malignancy4 %>% rename(patients = Mixture)

stats_Malignancy4 <- Malignancy4 %>%
  group_by(cell_types) %>%
  summarise(
    mean_fraction = mean(cell_fraction, na.rm = TRUE),
    sd_fraction = sd(cell_fraction, na.rm = TRUE)
  ) %>%
  mutate(
    log2_mean_fraction = log2(mean_fraction + 1),
    log2_sd_fraction = log2(sd_fraction + 1)
  )

M4 <- ggplot(stats_Malignancy4, aes(x = log2_mean_fraction, y = log2_sd_fraction, color = cell_types, label = cell_types)) +
  geom_point(size = 3) +
  geom_text(vjust = -0.5, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red",linetype = "dashed") +
  scale_color_manual(name = "Cell Types", values = stackcolors) +
  coord_cartesian(xlim = c(-0.15, 0.6), ylim = c(0, 0.15)) +
  theme_classic() +
  labs(
    title = "Cell type representation of patients with Lehismaniosis",
    x = "Log2(Mean Cell Fraction +1)",
    y = "Log2(Standard Deviation +1)",
    caption = "(Based on LM22 Signature Matrix)"
  ) +
  theme(
    legend.text = element_text(size = 5),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.3, "cm"),
    plot.title = element_text(size = 10)
  )
print(M4)
```

# VdW VS Different conditions
Starting with Malignancy 1 of Brucellosis patients with complete remission after treatment
```{r}
Malignancy1_VdW <- read.csv("CIBERSORTx_Results.csv")

Malignancy1_VdW <- Malignancy1_VdW[, 1:(ncol(Malignancy1_VdW)-3)]  

Malignancy1_VdW <- Malignancy1_VdW %>% pivot_longer(!Mixture, names_to = "cell_types", values_to = "cell_fraction")

Malignancy1_VdW <- Malignancy1_VdW %>% rename(patients = Mixture)

stats_Malignancy1_VdW <- Malignancy1_VdW %>%
  group_by(cell_types) %>%
  summarise(
    mean_fraction = mean(cell_fraction, na.rm = TRUE),
    sd_fraction = sd(cell_fraction, na.rm = TRUE)
  ) %>%
  mutate(
    log2_mean_fraction = log2(mean_fraction + 1),
    log2_sd_fraction = log2(sd_fraction + 1)
  )

M1_VdW <- ggplot(stats_Malignancy1_VdW, aes(x = log2_mean_fraction, y = log2_sd_fraction, color = cell_types, label = cell_types)) +
  geom_point(size = 3) +
  geom_text(vjust = -0.5, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red",linetype = "dashed") +
  scale_color_manual(name = "Cell Types", values = stackcolors2) +
  coord_cartesian(xlim = c(-0.03, 0.6), ylim = c(0, 0.35)) +
  theme_classic() +
  labs(
    title = "Cell type representation of Brucellosis patients with complete remission after treatment",
    x = "Log2(Mean Cell Fraction +1)",
    y = "Log2(Standard Deviation +1)",
    caption = "(Based on Van der Wijst Signature Matrix)"
  ) +
  theme(
    legend.text = element_text(size = 5),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.3, "cm"),
    plot.title = element_text(size = 10)
  )
print(M1_VdW)
```

Now Malignancy 2 wich is Brucellosis patients with reinfection after treatment
```{r}
Malignancy2_VdW <- read.csv("CIBERSORTx_Results.csv")

Malignancy2_VdW <- Malignancy2_VdW[, 1:(ncol(Malignancy2_VdW)-3)]  

Malignancy2_VdW <- Malignancy2_VdW %>% pivot_longer(!Mixture, names_to = "cell_types", values_to = "cell_fraction")

Malignancy2_VdW <- Malignancy2_VdW %>% rename(patients = Mixture)

stats_Malignancy2_VdW <- Malignancy2_VdW %>%
  group_by(cell_types) %>%
  summarise(
    mean_fraction = mean(cell_fraction, na.rm = TRUE),
    sd_fraction = sd(cell_fraction, na.rm = TRUE)
  ) %>%
  mutate(
    log2_mean_fraction = log2(mean_fraction + 1),
    log2_sd_fraction = log2(sd_fraction + 1)
  )

M2_VdW <- ggplot(stats_Malignancy2_VdW, aes(x = log2_mean_fraction, y = log2_sd_fraction, color = cell_types, label = cell_types)) +
  geom_point(size = 3) +
  geom_text(vjust = -0.5, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red",linetype = "dashed") +
  scale_color_manual(name = "Cell Types", values = stackcolors2) +
  coord_cartesian(xlim = c(-0.03, 0.6), ylim = c(0, 0.35)) +
  theme_classic() +
  labs(
    title = "Cell type representation of Brucellosis patients with reinfection after treatment",
    x = "Log2(Mean Cell Fraction +1)",
    y = "Log2(Standard Deviation +1)",
    caption = "(Based on Van der Wijst Signature Matrix)"
  ) +
  theme(
    legend.text = element_text(size = 5),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.3, "cm"),
    plot.title = element_text(size = 10)
  )
print(M2_VdW)
```

Now Malignancy 3 wich is Brucellosis patients with relapse after treatment
```{r}
Malignancy3_VdW <- read.csv("CIBERSORTx_Results.csv")

Malignancy3_VdW <- Malignancy3_VdW[, 1:(ncol(Malignancy3_VdW)-3)]  

Malignancy3_VdW <- Malignancy3_VdW %>% pivot_longer(!Mixture, names_to = "cell_types", values_to = "cell_fraction")

Malignancy3_VdW <- Malignancy3_VdW %>% rename(patients = Mixture)

stats_Malignancy3_VdW <- Malignancy3_VdW %>%
  group_by(cell_types) %>%
  summarise(
    mean_fraction = mean(cell_fraction, na.rm = TRUE),
    sd_fraction = sd(cell_fraction, na.rm = TRUE)
  ) %>%
  mutate(
    log2_mean_fraction = log2(mean_fraction + 1),
    log2_sd_fraction = log2(sd_fraction + 1)
  )

M3_VdW <- ggplot(stats_Malignancy3_VdW, aes(x = log2_mean_fraction, y = log2_sd_fraction, color = cell_types, label = cell_types)) +
  geom_point(size = 3) +
  geom_text(vjust = -0.5, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red",linetype = "dashed") +
  scale_color_manual(name = "Cell Types", values = stackcolors2) +
  coord_cartesian(xlim = c(-0.03, 0.6), ylim = c(0, 0.35)) +
  theme_classic() +
  labs(
    title = "Cell type representation of Brucellosis patients with relapse after treatment",
    x = "Log2(Mean Cell Fraction +1)",
    y = "Log2(Standard Deviation +1)",
    caption = "(Based on Van der Wijst)"
  ) +
  theme(
    legend.text = element_text(size = 5),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.3, "cm"),
    plot.title = element_text(size = 10)
  )
print(M3_VdW)
```

Now Malignancy 4 wich is Lehismaniosis patients
```{r}
Malignancy4_VdW <- read.csv("CIBERSORTx_Results.csv")

Malignancy4_VdW <- Malignancy4_VdW[, 1:(ncol(Malignancy4_VdW)-3)]  
Malignancy4_VdW <- Malignancy4_VdW %>% pivot_longer(!Mixture, names_to = "cell_types", values_to = "cell_fraction")
Malignancy4_VdW <- Malignancy4_VdW %>% rename(patients = Mixture)

stats_Malignancy4_VdW <- Malignancy4_VdW %>%
  group_by(cell_types) %>%
  summarise(
    mean_fraction = mean(cell_fraction, na.rm = TRUE),
    sd_fraction = sd(cell_fraction, na.rm = TRUE)
  ) %>%
  mutate(
    log2_mean_fraction = log2(mean_fraction + 1),
    log2_sd_fraction = log2(sd_fraction + 1)
  )

M4_VdW <- ggplot(stats_Malignancy4_VdW, aes(x = log2_mean_fraction, y = log2_sd_fraction, color = cell_types, label = cell_types)) +
  geom_point(size = 3) +
  geom_text(vjust = -0.5, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red",linetype = "dashed") +
  scale_color_manual(name = "Cell Types", values = stackcolors2) +
  coord_cartesian(xlim = c(-0.03, 0.6), ylim = c(0, 0.35)) +
  theme_classic() +
  labs(
    title = "Cell type representation of patients with Leishmaniasis",
    x = "Log2(Mean Cell Fraction +1)",
    y = "Log2(Standard Deviation +1)",
    caption = "(Based on Van der Wijst)"
  ) +
  theme(
    legend.text = element_text(size = 5),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.3, "cm"),
    plot.title = element_text(size = 10)
  )
print(M4_VdW)
```

# Default SM from ABIS  VS Different conditions
Malignancy 1
```{r}
Malignancy1_ABIS <- read.csv("CIBERSORTx_Results.csv")

Malignancy1_ABIS <- Malignancy1_ABIS[, 1:(ncol(Malignancy1_ABIS)-3)]  

Malignancy1_ABIS <- Malignancy1_ABIS %>% pivot_longer(!Mixture, names_to = "cell_types", values_to = "cell_fraction")

Malignancy1_ABIS <- Malignancy1_ABIS %>% rename(patients = Mixture)

stats_Malignancy1_ABIS <- Malignancy1_ABIS %>%
  group_by(cell_types) %>%
  summarise(
    mean_fraction = mean(cell_fraction, na.rm = TRUE),
    sd_fraction = sd(cell_fraction, na.rm = TRUE)
  ) %>%
  mutate(
    log2_mean_fraction = log2(mean_fraction + 1),
    log2_sd_fraction = log2(sd_fraction + 1)
  )

M1_ABIS <- ggplot(stats_Malignancy1_ABIS, aes(x = log2_mean_fraction, y = log2_sd_fraction, color = cell_types, label = cell_types)) +
  geom_point(size = 3) +
  geom_text(vjust = -0.5, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red",linetype = "dashed") +
  scale_color_manual(name = "Cell Types", values = stackcolors3) +
  coord_cartesian(xlim = c(-0.03, 0.6), ylim = c(0, 0.20)) +
  theme_classic() +
  labs(
    title = "Cell type representation of Brucellosis patients with complete remission after treatment",
    x = "Log2(Mean Cell Fraction +1)",
    y = "Log2(Standard Deviation +1)",
    caption = "(Based on ABIS default SM)"
  ) +
  theme(
    legend.text = element_text(size = 5),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.3, "cm"),
    plot.title = element_text(size = 10)
  )
print(M1_ABIS)
```

Malignancy 2
```{r}
Malignancy2_ABIS <- read.csv("CIBERSORTx_Results.csv")

Malignancy2_ABIS <- Malignancy2_ABIS[, 1:(ncol(Malignancy2_ABIS)-3)]  
Malignancy2_ABIS <- Malignancy2_ABIS %>% pivot_longer(!Mixture, names_to = "cell_types", values_to = "cell_fraction")
Malignancy2_ABIS <- Malignancy2_ABIS %>% rename(patients = Mixture)

stats_Malignancy2_ABIS <- Malignancy2_ABIS %>%
  group_by(cell_types) %>%
  summarise(
    mean_fraction = mean(cell_fraction, na.rm = TRUE),
    sd_fraction = sd(cell_fraction, na.rm = TRUE)
  ) %>%
  mutate(
    log2_mean_fraction = log2(mean_fraction + 1),
    log2_sd_fraction = log2(sd_fraction + 1)
  )

M2_ABIS <- ggplot(stats_Malignancy2_ABIS, aes(x = log2_mean_fraction, y = log2_sd_fraction, color = cell_types, label = cell_types)) +
  geom_point(size = 3) +
  geom_text(vjust = -0.5, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red",linetype = "dashed") +
  scale_color_manual(name = "Cell Types", values = stackcolors3) +
  coord_cartesian(xlim = c(-0.03, 0.6), ylim = c(0, 0.20)) +
  theme_classic() +
  labs(
    title = "Cell type representation of Brucellosis patients with reinfection after treatment",
    x = "Log2(Mean Cell Fraction +1)",
    y = "Log2(Standard Deviation +1)",
    caption = "(Based on ABIS default SM)"
  ) +
  theme(
    legend.text = element_text(size = 5),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.3, "cm"),
    plot.title = element_text(size = 10)
  )
print(M2_ABIS)
```

Malignancy 3
```{r}
Malignancy3_ABIS <- read.csv("CIBERSORTx_Results.csv")

Malignancy3_ABIS <- Malignancy3_ABIS[, 1:(ncol(Malignancy3_ABIS)-3)]  
Malignancy3_ABIS <- Malignancy3_ABIS %>% pivot_longer(!Mixture, names_to = "cell_types", values_to = "cell_fraction")
Malignancy3_ABIS <- Malignancy3_ABIS %>% rename(patients = Mixture)

stats_Malignancy3_ABIS <- Malignancy3_ABIS %>%
  group_by(cell_types) %>%
  summarise(
    mean_fraction = mean(cell_fraction, na.rm = TRUE),
    sd_fraction = sd(cell_fraction, na.rm = TRUE)
  ) %>%
  mutate(
    log2_mean_fraction = log2(mean_fraction + 1),
    log2_sd_fraction = log2(sd_fraction + 1)
  )

M3_ABIS <- ggplot(stats_Malignancy3_ABIS, aes(x = log2_mean_fraction, y = log2_sd_fraction, color = cell_types, label = cell_types)) +
  geom_point(size = 3) +
  geom_text(vjust = -0.5, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red",linetype = "dashed") +
  scale_color_manual(name = "Cell Types", values = stackcolors3) +
  coord_cartesian(xlim = c(-0.03, 0.6), ylim = c(0, 0.20)) +
  theme_classic() +
  labs(
    title = "Cell type representation of Brucellosis patients with relapse after treatment",
    x = "Log2(Mean Cell Fraction +1)",
    y = "Log2(Standard Deviation +1)",
    caption = "(Based on ABIS default SM)"
  ) +
  theme(
    legend.text = element_text(size = 5),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.3, "cm"),
    plot.title = element_text(size = 10)
  )
print(M3_ABIS)
```

Malignancy 4
```{r}
Malignancy4_ABIS <- read.csv("CIBERSORTx_Results.csv")

Malignancy4_ABIS <- Malignancy4_ABIS[, 1:(ncol(Malignancy4_ABIS)-3)]  
Malignancy4_ABIS <- Malignancy4_ABIS %>% pivot_longer(!Mixture, names_to = "cell_types", values_to = "cell_fraction")
Malignancy4_ABIS <- Malignancy4_ABIS %>% rename(patients = Mixture)

stats_Malignancy4_ABIS <- Malignancy4_ABIS %>%
  group_by(cell_types) %>%
  summarise(
    mean_fraction = mean(cell_fraction, na.rm = TRUE),
    sd_fraction = sd(cell_fraction, na.rm = TRUE)
  ) %>%
  mutate(
    log2_mean_fraction = log2(mean_fraction + 1),
    log2_sd_fraction = log2(sd_fraction + 1)
  )

M4_ABIS <- ggplot(stats_Malignancy4_ABIS, aes(x = log2_mean_fraction, y = log2_sd_fraction, color = cell_types, label = cell_types)) +
  geom_point(size = 3) +
  geom_text(vjust = -0.5, size = 3) +
  geom_smooth(method = "lm", se = FALSE, color = "red",linetype = "dashed") +
  scale_color_manual(name = "Cell Types", values = stackcolors3) +
  coord_cartesian(xlim = c(-0.03, 0.6), ylim = c(0, 0.20)) +
  theme_classic() +
  labs(
    title = "Cell type representation of patients with Lehismaniosis",
    x = "Log2(Mean Cell Fraction +1)",
    y = "Log2(Standard Deviation +1)",
    caption = "(Based on ABIS default SM)"
  ) +
  theme(
    legend.text = element_text(size = 5),
    legend.title = element_text(size = 9),
    legend.key.size = unit(0.3, "cm"),
    plot.title = element_text(size = 10)
  )
print(M4_ABIS)
```

# Doing bar plot for Mean cell fractions and for Standard deviation fractions for selected cell types*

*First for LM22 SM*
```{r}
d1 <- stats_TM1_LM22.log[stats_TM1_LM22.log$cell_types %in% c(	
"Neutrophils","T.cells.CD8","NK.cells.resting",	
"T.cells.CD4.naive","Monocytes", "T.cells.CD4.memory.resting"),c(1,4,5)]
d1$Condition <- "Healthy"

d2 <- stats_Malignancy1[stats_Malignancy1$cell_types %in% c(	
"Neutrophils","T.cells.CD8","NK.cells.resting",	
"T.cells.CD4.naive","Monocytes", "T.cells.CD4.memory.resting"
),c(1,4,5)]
d2$Condition <- "Brucellosis patients with complete remission after treatment"

d3 <- stats_Malignancy2[stats_Malignancy2$cell_types %in% c(	
"Neutrophils","T.cells.CD8","NK.cells.resting",	
"T.cells.CD4.naive","Monocytes", "T.cells.CD4.memory.resting"),c(1,4,5)]
d3$Condition <- "Brucellosis patients with reinfection after treatment"

d4 <- stats_Malignancy3[stats_Malignancy3$cell_types %in% c(	
"Neutrophils","T.cells.CD8","NK.cells.resting",	
"T.cells.CD4.naive","Monocytes", "T.cells.CD4.memory.resting"),c(1,4,5)]
d4$Condition <- "Brucellosis patients with relapse after treatment"

d5 <- stats_Malignancy4[stats_Malignancy4$cell_types %in% c(	
"Neutrophils","T.cells.CD8","NK.cells.resting",	
"T.cells.CD4.naive","Monocytes", "T.cells.CD4.memory.resting"),c(1,4,5)]
d5$Condition <- "Patients with Leishmaniasis"

LM22_study <- rbind(d1, d2, d3, d4, d5) 
colnames(LM22_study)[colnames(LM22_study) == "log2_mean_fraction"] <- "MeanFraction"
colnames(LM22_study)[colnames(LM22_study) == "cell_types"] <- "CellType"

LM22_study$Condition <- factor(LM22_study$Condition, levels = c(
  "Healthy",
  "Brucellosis patients with complete remission after treatment",
  "Brucellosis patients with reinfection after treatment",
  "Brucellosis patients with relapse after treatment",
  "Patients with Leishmaniasis"
))

s1 <- ggplot(LM22_study, aes(x = CellType, y = MeanFraction, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(x = "Cell type", y = "log2(Mean_Fraction+1)") +
  theme_minimal() +
  scale_fill_manual(values = c(
    "Healthy" = "grey", 
    "Brucellosis patients with complete remission after treatment" = "green", 
    "Brucellosis patients with reinfection after treatment" = "darkgreen",
    "Brucellosis patients with relapse after treatment" = "red",
    "Patients with Leishmaniasis" = "gold"
  )) +
  scale_y_continuous(breaks = seq(0, 0.6, by = 0.05)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(s1)
```
Standard deviation values for LM22
```{r}
colnames(LM22_study)[colnames(LM22_study) == "log2_sd_fraction"] <- "SDFraction"
s2 <-ggplot(LM22_study, aes(x = CellType, y = SDFraction, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(x = "Cell type", y = "log2(SD_Fraction+1)") +
  theme_minimal() +
    scale_fill_manual(values = c("Healthy" = "grey", 
                             "Brucellosis patients with complete remission after treatment" = "green", 
                             "Brucellosis patients with reinfection after treatment" = "darkgreen",
                             "Brucellosis patients with relapse after treatment" = "red",
                             "Patients with Leishmaniasis" = "gold")) +
  scale_y_continuous(breaks = seq(0, 0.2, by = 0.02)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
print(s2)
```

*Doing the same for Van der Wijst*
```{r}
d6 <- stats_TM1_VdW.log[stats_TM1_VdW.log$cell_types %in% c(	
"CD8_T", "CD4_T", "CD56dim_NK",	"cMonocyte","Plasma","B"),c(1,4,5)]
d6$Condition <- "Healthy"

d7 <- stats_Malignancy1_VdW[stats_Malignancy1_VdW$cell_types %in% c(	
"CD8_T", "CD4_T", "CD56dim_NK",	"cMonocyte","Plasma","B"),c(1,4,5)]
d7$Condition <- "Brucellosis patients with complete remission after treatment"

d8 <- stats_Malignancy2_VdW[stats_Malignancy2_VdW$cell_types %in% c(	
"CD8_T", "CD4_T", "CD56dim_NK",	"cMonocyte","Plasma","B"),c(1,4,5)]
d8$Condition <- "Brucellosis patients with reinfection after treatment"

d9 <- stats_Malignancy3_VdW[stats_Malignancy3_VdW$cell_types %in% c(	
"CD8_T", "CD4_T", "CD56dim_NK",	"cMonocyte","Plasma","B"),c(1,4,5)]
d9$Condition <- "Brucellosis patients with relapse after treatment"

d10 <- stats_Malignancy4_VdW[stats_Malignancy4_VdW$cell_types %in% c(	
"CD8_T", "CD4_T", "CD56dim_NK",	"cMonocyte","Plasma","B"),c(1,4,5)]
d10$Condition <- "Patients with Leishmaniasis"

VdW_study <- rbind(d6, d7, d8, d9, d10)  
colnames(VdW_study)[colnames(VdW_study) == "log2_mean_fraction"] <- "MeanFraction"
colnames(VdW_study)[colnames(VdW_study) == "cell_types"] <- "CellType"
VdW_study$Condition <- factor(LM22_study$Condition, levels = c(
  "Healthy",
  "Brucellosis patients with complete remission after treatment",
  "Brucellosis patients with reinfection after treatment",
  "Brucellosis patients with relapse after treatment",
  "Patients with Leishmaniasis"
))

s3 <-ggplot(VdW_study, aes(x = CellType, y = MeanFraction, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(x = "Cell type", y = "log2(Mean_Fraction+1)") +
  theme_minimal() +
  scale_fill_manual(values = c("Healthy" = "grey", 
                             "Brucellosis patients with complete remission after treatment" = "green", 
                             "Brucellosis patients with reinfection after treatment" = "darkgreen",
                             "Brucellosis patients with relapse after treatment" = "red",
                             "Patients with Leishmaniasis" = "gold")) +
  scale_y_continuous(breaks = seq(0, 0.6, by = 0.05)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(s3)
```

```{r}
colnames(VdW_study)[colnames(VdW_study) == "log2_sd_fraction"] <- "SDFraction"
s4 <-ggplot(VdW_study, aes(x = CellType, y = SDFraction, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(x = "Cell type", y = "log2(SD_Fraction+1)") +
  theme_minimal() +
    scale_fill_manual(values = c("Healthy" = "grey", 
                             "Brucellosis patients with complete remission after treatment" = "green", 
                             "Brucellosis patients with reinfection after treatment" = "darkgreen",
                             "Brucellosis patients with relapse after treatment" = "red",
                             "Patients with Leishmaniasis" = "gold")) +
  scale_y_continuous(breaks = seq(0, 0.4, by = 0.05)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
print(s4)
```

*Also for Default SM from ABIS*

Now the bar plot comparision of the Means and Standard deviation to see if there is any difference
```{r}
d11 <- stats_TM1_ABIS[stats_TM1_ABIS$cell_types %in% c(	
"Neutrophils.LD","T.CD8.Memory","T.CD4.Naive",	
"T.gd.non.Vd2"),c(1,4,5)]
d11$Condition <- "Healthy"

d12 <- stats_Malignancy1_ABIS[stats_Malignancy1_ABIS$cell_types %in% c(	
"Neutrophils.LD","T.CD8.Memory","T.CD4.Naive",	
"T.gd.non.Vd2"),c(1,4,5)]
d12$Condition <- "Brucellosis patients with complete remission after treatment"

d13 <- stats_Malignancy2_ABIS[stats_Malignancy2_ABIS$cell_types %in% c(	
"Neutrophils.LD","T.CD8.Memory","T.CD4.Naive",	
"T.gd.non.Vd2"),c(1,4,5)]
d13$Condition <- "Brucellosis patients with reinfection after treatment"

d14 <- stats_Malignancy3_ABIS[stats_Malignancy3_ABIS$cell_types %in% c(	
"Neutrophils.LD","T.CD8.Memory","T.CD4.Naive",	
"T.gd.non.Vd2"),c(1,4,5)]
d14$Condition <- "Brucellosis patients with relapse after treatment"


d15 <- stats_Malignancy4_ABIS[stats_Malignancy4_ABIS$cell_types %in% c(	
"Neutrophils.LD","T.CD8.Memory","T.CD4.Naive",	
"T.gd.non.Vd2"),c(1,4,5)]
d15$Condition <- "Patients with Lehismaniosis"

ABIS_Monaco_study <- rbind(d11, d12, d13, d14, d15) 

colnames(ABIS_Monaco_study)[colnames(ABIS_Monaco_study) == "log2_mean_fraction"] <- "MeanFraction"

colnames(ABIS_Monaco_study)[colnames(ABIS_Monaco_study) == "cell_types"] <- "CellType"


s5 <-ggplot(ABIS_Monaco_study, aes(x = CellType, y = MeanFraction, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(x = "Cell type", y = "log2(Mean_Fraction+1)") +
  theme_minimal() +
  scale_fill_manual(values = c("Healthy" = "grey", 
                             "Brucellosis patients with complete remission after treatment" = "green", 
                             "Brucellosis patients with reinfection after treatment" = "darkgreen",
                             "Brucellosis patients with relapse after treatment" = "red",
                             "Patients with Lehismaniosis" = "gold")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

print(s5)
```
```{r}
colnames(ABIS_Monaco_study)[colnames(ABIS_Monaco_study) == "log2_sd_fraction"] <- "SDFraction"
s6 <-ggplot(ABIS_Monaco_study, aes(x = CellType, y = SDFraction, fill = Condition)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  labs(x = "Cell type", y = "log2(SD_Fraction+1)") +
  theme_minimal() +
    scale_fill_manual(values = c("Healthy" = "grey", 
                             "Brucellosis patients with complete remission after treatment" = "green", 
                             "Brucellosis patients with reinfection after treatment" = "darkgreen",
                             "Brucellosis patients with relapse after treatment" = "red",
                             "Patients with Lehismaniosis" = "gold")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
print(s6)
```

# Calculating the % of difference between healthy CD4 T cells and CD8 T cells compared to different conditions (by Van der Wijst)

```{r}
# For healthy:
#CD8+T
print(stats_TM1_VdW.log[10,"mean_fraction"] *100)
#CD4+T
print(stats_TM1_VdW.log[11,"mean_fraction"] *100)

# Brucellosis with (malignancy 1)
#CD8+T
print(stats_Malignancy1_VdW[5,"mean_fraction"] *100)
#CD4+T
print(stats_Malignancy1_VdW[2,"mean_fraction"] *100)

# Brucellosis with (malignancy 2)
#CD8+T
print(stats_Malignancy2_VdW[5,"mean_fraction"] *100)
#CD4+T
print(stats_Malignancy2_VdW[2,"mean_fraction"] *100)

# Brucellosis with (malignancy 3)
#CD8+T
print(stats_Malignancy3_VdW[5,"mean_fraction"] *100)
#CD4+T
print(stats_Malignancy3_VdW[2,"mean_fraction"] *100)
# Leishmaniasis
#CD8+T
print(stats_Malignancy4_VdW[5,"mean_fraction"] *100)
#CD4+T
print(stats_Malignancy4_VdW[2,"mean_fraction"] *100)
```


