---
title: "Creation_of_diseases_datasets"
output: html_notebook
---
Document following the creation of the four different conditions

# Creating Four small datasets:
- Brucellosis (complete remission after treatment)  Name:Malignancy1
- Brucellosis (reinfection after treatment)         Name:Malignancy2
- Brucellosis (relapse after treatment)             Name:Malignancy3
- Lehismaniosis (condition not aviable)             Name:Malignancy4

# Libraries needed:
```{r}
library(Seurat)
library(dplyr)
library(tibble)
```

# Creating the datasets
```{r}
Brucellosis_ind <- c() 
Brucellosis_cond <- c() 
Brucellosis_stage <- c() 

for (a in 1:nrow(database_3_experimentdesign)) {
  if (grepl("brucellosis", database_3_experimentdesign[a, "Sample.Characteristic.disease."])) {
    Brucellosis_ind <- c(Brucellosis_ind, database_3_experimentdesign[a, "Run"])
    Brucellosis_cond <- c(Brucellosis_cond, database_3_experimentdesign[a, "Sample.Characteristic.clinical.information."])
    Brucellosis_stage <- c(Brucellosis_stage, database_3_experimentdesign[a, "Sample.Characteristic.disease.staging."])
  }
}
print(Brucellosis_ind)
print(Brucellosis_cond)

Brucellosis_patients <- data.frame(
  IndividualID = Brucellosis_ind,
  Condition = Brucellosis_cond,
  Stage = Brucellosis_stage
)
```

*Malignancy 1*
```{r}
Malignancy1 <- data.frame("Gene Symbol"=rawdatabase_3$Gene.Name)
for (a in 1:nrow(Brucellosis_patients)) { 
  patient <- Brucellosis_patients$IndividualID[a]
  matching_col <- which(colnames(rawdatabase_3) == patient) # This gives the number of column that the names are the same, if there isn't any match, gives 0
  condition <- Brucellosis_patients$Condition[a]
  if (condition == "complete remission after treatment" && length(matching_col) == 1) {  #
    Malignancy1[[patient]] <- rawdatabase_3[, matching_col]  # Add each column with its information
  }
}
```

*Malignancy 2*
```{r}
Malignancy2 <- data.frame("Gene Symbol"=rawdatabase_3$Gene.Name)
for (a in 1:nrow(Brucellosis_patients)) { 
  patient <- Brucellosis_patients$IndividualID[a]
  matching_col <- which(colnames(rawdatabase_3) == patient) # This gives the number of column that the names are the same, if there isn't any match, gives 0
  condition <- Brucellosis_patients$Condition[a]
  if (condition == "reinfection after treatment" && length(matching_col) == 1) {  #
    Malignancy2[[patient]] <- rawdatabase_3[, matching_col]  # Add each column with its information
  }
}
```

*Malignancy 3*
```{r}
Malignancy3 <- data.frame("Gene Symbol"=rawdatabase_3$Gene.Name)
for (a in 1:nrow(Brucellosis_patients)) { 
  patient <- Brucellosis_patients$IndividualID[a]
  matching_col <- which(colnames(rawdatabase_3) == patient) # This gives the number of column that the names are the same, if there isn't any match, gives 0
  condition <- Brucellosis_patients$Condition[a]
  if (condition == "relapse after treatment" && length(matching_col) == 1) {  #
    Malignancy3[[patient]] <- rawdatabase_3[, matching_col]  # Add each column with its information
  }
}
```

*Malignancy 4*
```{r}
Leishmaniasis_ind <- c() 

for (a in 1:nrow(database_3_experimentdesign)) {
  if (grepl("Leishmaniasis", database_3_experimentdesign[a, "Sample.Characteristic.disease."])) {
    Leishmaniasis_ind <- c(Leishmaniasis_ind, database_3_experimentdesign[a, "Run"])
  }
}
print(Leishmaniasis_ind)

Malignancy4 <- data.frame("Gene Symbol"=rawdatabase_3$Gene.Name)
for (a in 1:length(Leishmaniasis_ind)) { # Using length instead of ncol, as it isn't a data frame.
  patient <- Leishmaniasis_ind[a]
  matching_col <- which(colnames(rawdatabase_3) == patient) # This gives the number of column that the names are the same, if there isn't any match, gives 0
  if (length(matching_col) == 1) {  #
    Malignancy4[[patient]] <- rawdatabase_3[, matching_col]  # Add each column with its information
  }
}
```


# Calculating TPM
```{r}
TPM_calculator2 <- function(database,MANE) {
  
  for (c in 1:ncol(database)) {
    for(r in 1:nrow(database)) {
      gene_name <- rownames(database)[r]
      row_number <- match(gene_name, MANE$Gene.name)
      RPKpergene <- database[r,c]/(MANE[row_number,"MANE_length"]/1000)
      database[r,c] <- RPKpergene
    }  
  }

  database <- na.omit(database)
  sum_RPKs_per_sample <- colSums(database)
  database <- sweep(database, 2, sum_RPKs_per_sample, "/") * 1e6
  print(colSums(database))
  database<- round(database, 2)
  return(database)
}
MANE <- way3_MANE_data_select_transcript_length
MANE <- MANE %>%
  left_join(
    select(Biomart_information, Transcript.stable.ID, Gene.name),
    by = c("Transcript_ID" = "Transcript.stable.ID")
  )

# Malignancy1 TPMs
Malignancy1 <- Malignancy1 %>%
  filter(!is.na(Gene.Symbol) & trimws(Gene.Symbol) != "") %>%
  group_by(Gene.Symbol) %>%
  slice_max(order_by = rowSums(across(where(is.numeric), ~ . > 1)), n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  distinct(Gene.Symbol, .keep_all = TRUE)

Malignancy1 <- as.data.frame(Malignancy1)
rownames(Malignancy1) <- Malignancy1$Gene.Symbol
Malignancy1$Gene.Symbol <- NULL
Malignancy1 <- TPM_calculator2(Malignancy1, MANE)

# Malignancy 2 TPMs
Malignancy2 <- Malignancy2 %>%
  filter(!is.na(Gene.Symbol) & trimws(Gene.Symbol) != "") %>%
  group_by(Gene.Symbol) %>%
  slice_max(order_by = rowSums(across(where(is.numeric), ~ . > 1)), n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  distinct(Gene.Symbol, .keep_all = TRUE)

Malignancy2 <- as.data.frame(Malignancy2)
rownames(Malignancy2) <- Malignancy2$Gene.Symbol
Malignancy2$Gene.Symbol <- NULL
Malignancy2 <- TPM_calculator2(Malignancy2, MANE)

# Malignancy 3 TPMs
Malignancy3 <- Malignancy3 %>%
  filter(!is.na(Gene.Symbol) & trimws(Gene.Symbol) != "") %>%
  group_by(Gene.Symbol) %>%
  slice_max(order_by = rowSums(across(where(is.numeric), ~ . > 1)), n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  distinct(Gene.Symbol, .keep_all = TRUE)

Malignancy3 <- as.data.frame(Malignancy3)
rownames(Malignancy3) <- Malignancy3$Gene.Symbol
Malignancy3$Gene.Symbol <- NULL
Malignancy3 <- TPM_calculator2(Malignancy3, MANE)


# Malignancy 4 TPMs
Malignancy4 <- Malignancy4 %>%
  filter(!is.na(Gene.Symbol) & trimws(Gene.Symbol) != "") %>%
  group_by(Gene.Symbol) %>%
  slice_max(order_by = rowSums(across(where(is.numeric), ~ . > 1)), n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  distinct(Gene.Symbol, .keep_all = TRUE)

Malignancy4 <- as.data.frame(Malignancy4)
rownames(Malignancy4) <- Malignancy4$Gene.Symbol
Malignancy4$Gene.Symbol <- NULL
Malignancy4 <- TPM_calculator2(Malignancy4, MANE)
```


# Preparing the datasets for CIBERSORTx
```{r}
Malignancy1 <-rownames_to_column(Malignancy1,var="Gene Symbol") 
Malignancy1 <- rbind(colnames(Malignancy1), Malignancy1) 
colnames(Malignancy1)<-NULL

Malignancy2 <-rownames_to_column(Malignancy2,var="Gene Symbol") 
Malignancy2 <- rbind(colnames(Malignancy2), Malignancy2) 
colnames(Malignancy2)<-NULL

Malignancy3 <-rownames_to_column(Malignancy3,var="Gene Symbol") 
Malignancy3 <- rbind(colnames(Malignancy3), Malignancy3) 
colnames(Malignancy3)<-NULL

Malignancy4 <-rownames_to_column(Malignancy4,var="Gene Symbol") 
Malignancy4 <- rbind(colnames(Malignancy4), Malignancy4) 
colnames(Malignancy4)<-NULL

write.table(Malignancy1, "Malignancy_1.txt", sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)
write.table(Malignancy2, "Malignancy_2.txt", sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)
write.table(Malignancy3, "Malignancy_3.txt", sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)
write.table(Malignancy4, "Malignancy_4.txt", sep = "\t", row.names = FALSE, col.names = FALSE, quote = FALSE)
```

